Estimating cardiac contraction through high resolution data assimilation of a personalized mechanical model

### 2017-09-26 chapters/introduction.tex
Added paper 3 and update introduction

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THE PAST %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Cardiovascular diseases are the leading cause of death in the
world, responsible for more than 17.3 million deaths annually, a
number that is expected to reach 23.6 million by 2030
\cite{writing2016heart}. Advancements in detection and treatment
have reduced morbidity and mortality, but knowledge of the underlying
mechanisms for the development of the disease is still very poor.
While some patients respond very well to a therapy, others suffering
from the same symptoms might not respond at all.
Our understanding of these mechanisms has been improved through
development of advance imaging techniques and through
computational modeling. While imaging play a central role in
assessment of cardiac structure and kinematics, biophysical models can
help us to understand causality in this multiscale environment.

@@

One of the earliest examples of modelling applied to cardiac
physiology dates back to 1892 when Robert Wood approximated the
ventricle as a thinned-wall spherical shell \cite{woods1892few}, and
formulated the well known \emph{Law of Laplace}.
% In this way, \emph{wall
% stress} could be analytically estimated via the famous \emph{Law of Laplace}
% \cite{woods1892few}. This law relates the wall stress $\sigma$ to the endocardial
% pressure $P$, the endocarial radius $r$ and the wall thickness $w$
% and can in its simplest for be expressed as
% \begin{align}
%   \sigma = \frac{P \times r}{2 w}.
% \end{align}
% Wall stress can be thought of as the load that the heart must work
% against, and it is therefore important to understand how different
% quantities affects the working heart. For instance when the heart
% ejects blood, the load that the heart ejects blood against is
% typically referred to as \emph{afterload}, but this is really the same
% as the wall stress in that particular state. 
% Even though it is obvious that the shape of a ventricle in most cases
% clearly deviates from a sphere, this forumla provides clinician with a tool
% that can be used to understand how diseases affect the heart. Take for
% example the case of ventricular dilation, when the inner radius of 
% the ventricle increases, resulting in increased wall stress and
% hence the heart has to work harder in order to eject blood to the body. In the other
% end we have the case of concentric hypertrophy which results in wall
% thickening and hence reduced the wall stress. In this case more muscle
% fibers are recruited and therefore the tension experienced by each
% individual unit will be less. 
Since 1892 more realistic and sophisticated models have been developed
in order to better describe the anatomical as well as the biophysical
nature of the ventricle. The first extension came in 1927 to
thick-walled spheres \cite{love2013treatise}. In the 1960s the spheres
were substituted by ellipsoids, with thin-walled models in 1963
\cite{sandler1963left}, and with thick-walled models in 1968
\cite{wong1968stress}. A thorough comparison of the different early
models  is presented in \cite{huisman1980comparison}. The first
reported usage of image-based geometries dates back to 1972
\cite{gould1972vivo}, were X-ray was used to image the left ventricle
and the finite element method was used to estimate wall stress. The relatively
poor computational power, together with the low image quality really
challenged the biomedical bioengineers at that time. The following
phrase, taken form \cite{heethaar1977computer} illustrates this fact:  
\emph{``Programming for the solution of the large number of
  simultaneous linear equations generated by the finite element method
  on computers with capacity not exceeding 32K words of memory
  required special attention to very compact storage of the
  stiffness matrix and the retrieval of its coefficients''}.
A paradigm shift in the field of cardiac mechanical modeling came in
the late 1980s at the University of Auckland. Here the first models
based on a continuum description of the heart were developed
\cite{hunter1988analysis}, which is still in use today.

@@

These mechanical models contain model parameters which may vary from
individual to individual, and patient specificity can be obtained by
fitting these model parameters to additional clinical data. This
process is often referred to as data assimilation, a field that has
its root in meteorology where one want to make a weather forecast based
on observations from strategically placed weather stations. The idea
is to construct a functional which represent the mismatch between the
model and the observations, and by changing the model parameters we
also change the value of this mismatch functional. Now, what we want to do is to
minimize this functional using these model parameters as control. At
the same time, the governing PDE, which depend upon these model
parameters, has to remain true. We therefore often
refer to such problems as PDE-constrained optimization problems.  
Solving such problems is challenging because evaluation of the
mismatch functional requires the solution of a nonlinear PDE, which is
computationally expensive, and we therefore want to keep the number of
functional evaluations at a minimum. The best way of doing this is
traverse the parameter space along the gradient descent, which is
guaranteed to bring us towards a local minimum. However, this requires
calculation of the functional gradient, or the derivative of the
mismatch functional with respect to the model parameters. As we will
see, the functional gradient can be computed efficiently by
considering the adjoint equation, and recent development in numerical
software tools \cite{farrell2013automated} allow us to exploit this
also numerically.

>>>>

\section{Aim of thesis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%% THE PAST %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Cardiovascular diseases are the leading cause of death in the
world, responsible for more than 17.3 million deaths annually, a
number that is expected to reach 23.6 million by 2030
\cite{writing2016heart}. Advancements in detection and treatment
have reduced morbidity and mortality, but knowledge of the underlying
mechanisms for the development of the disease is still very poor.
While some patients respond very well to a therapy, others suffering
from the same symptoms might not respond at all.
Our understanding of these mechanisms has been improved through
development of advance imaging techniques and through
computational modeling. While imaging play a central role in
assessment of cardiac structure and kinematics, biophysical models can
help us to understand causality in this multiscale environment.

@@

One of the earliest examples of modelling applied to cardiac
physiology dates back to 1892 when Robert Wood approximated the
ventricle as a thinned-wall spherical shell \cite{woods1892few}, and
formulated the well known \emph{Law of Laplace}.
% In this way, \emph{wall
% stress} could be analytically estimated via the famous \emph{Law of Laplace}
% \cite{woods1892few}. This law relates the wall stress $\sigma$ to the endocardial
% pressure $P$, the endocarial radius $r$ and the wall thickness $w$
% and can in its simplest for be expressed as
% \begin{align}
%   \sigma = \frac{P \times r}{2 w}.
% \end{align}
% Wall stress can be thought of as the load that the heart must work
% against, and it is therefore important to understand how different
% quantities affects the working heart. For instance when the heart
% ejects blood, the load that the heart ejects blood against is
% typically referred to as \emph{afterload}, but this is really the same
% as the wall stress in that particular state. 
% Even though it is obvious that the shape of a ventricle in most cases
% clearly deviates from a sphere, this forumla provides clinician with a tool
% that can be used to understand how diseases affect the heart. Take for
% example the case of ventricular dilation, when the inner radius of 
% the ventricle increases, resulting in increased wall stress and
% hence the heart has to work harder in order to eject blood to the body. In the other
% end we have the case of concentric hypertrophy which results in wall
% thickening and hence reduced the wall stress. In this case more muscle
% fibers are recruited and therefore the tension experienced by each
% individual unit will be less. 
Since 1892 more realistic and sophisticated models have been developed
in order to better describe the anatomical as well as the biophysical
nature of the ventricle. The first extension came in 1927 to
thick-walled spheres \cite{love2013treatise}. In the 1960s the spheres
were substituted by ellipsoids, with thin-walled models in 1963
\cite{sandler1963left}, and with thick-walled models in 1968
\cite{wong1968stress}. A thorough comparison of the different early
models  is presented in \cite{huisman1980comparison}. The first
reported usage of image-based geometries dates back to 1972
\cite{gould1972vivo}, were X-ray was used to image the left ventricle
and the finite element method was used to estimate wall stress. The relatively
poor computational power, together with the low image quality really
challenged the biomedical bioengineers at that time. The following
phrase, taken form \cite{heethaar1977computer} illustrates this fact:  
\emph{``Programming for the solution of the large number of
  simultaneous linear equations generated by the finite element method
  on computers with capacity not exceeding 32K words of memory
  required special attention to very compact storage of the
  stiffness matrix and the retrieval of its coefficients''}.

A paradigm shift in the field of cardiac mechanics modeling came in
the late 1980s at the University of Auckland. Here the first models
based on a continuum description of the heart were developed
\cite{hunter1988analysis}, which is still in use today.

@@

These mechanical models contain model parameters which may vary from
individual to individual, and patient specificity can be obtained by
fitting these model parameters to additional clinical data. This
process is often referred to as data assimilation, a field that has
its root in meteorology where one want to make a weather forecast based
on observations from strategically placed weather stations. The idea
is to construct a functional which represent the mismatch between the
model and the observations, and by changing the model parameters we
also change the value of this mismatch functional. Now, what we want to do is to
minimize this functional using these model parameters as control. At
the same time, the governing PDE, which depend upon these model
parameters, has to remain true. We therefore often
refer to such problems as PDE-constrained optimization problems.  
Solving such problems is challenging because evaluation of the
mismatch functional requires the solution of a nonlinear PDE, which is
computationally expensive, and we therefore want to keep the number of
functional evaluations at a minimum. A good way of doing this is
traverse the parameter space along the gradient descent, which is
guaranteed to bring us towards a local minimum. However, this requires
calculation of the functional gradient, or the derivative of the
mismatch functional with respect to the model parameters. As we will
see, the functional gradient can be computed efficiently by
considering the adjoint equation, and recent development in numerical
software tools \cite{farrell2013automated} allow us to exploit this
also numerically.

The developed framework in this thesis could be applied directly to
estimate high dimensional model parameters without any manual tuning of
parameters. The aim has been to develop a framework for automatic
generation of patient-specific simulations of the heart, by fitting
model-parameters to potentially ``any'' type of clinical data. In particular,
incorporating regional motion data to estimate spatially resolved
parameters has been an essential contribution in this thesis. This
type of automated simulation pipeline could potentially be integrated
as a part of a clinical diagnostic toolbox and be used to guide clinicians in the
decision-making and treatment planning. In particular, such models can
be used to assess information that is difficult or even impossible to
measure in the human heart, such as indices of contractility,
myocardial work and myofiber stress. While the first part of the
thesis (Paper 1) is concerned with developing a framework for
constraining a mechanics model to clinical data, the second part
(Paper 2, 3 and 4) evaluates potentially useful biomarkers extracted from
the personalized mechanics model. Future work should be geared towards
validation of such models, in order to unleash its potential in the
clinic.

### 2017-09-26 chapters/introduction/summary.tex
Added paper 3 and update introduction

In this paper we develop and test the pipeline for constructing a
patient specific mechanical simulation of a patient's heart, based on
clinical measurments and adjoint-based data assimilation
techniques.

@@

In this paper we extend the work in the two previous paper to
bi-ventricular geomtries, and use it to estimate patient-specific
myofiber stress and indices of contractility. Furthermore, we investigate
the senstivtiy of these computed features the helical fiber angle and
the choice of active modeling framework.

In this work we estimate the linear isotopic material parameter in
\eqref{eq:holzapel_trans}, spatially resolved on the LV and RV, by
minimizing the error in end-diastolic LV and RV cavity volumes. We also
apply a simplified algorithm for estimating in unloaded geometry, in
order to overcome issues related to estimating an unloaded
BiV-geometry, which is not well-posed. The amount of active
contraction is estimated during the active phase, by minimizing
simulated and meausured volumes and circumferential strain. The active
control parameter, which are $\gamma$ in
\eqref{eq:intro_active_strain_Fa_gjerald}  and $T_a$ in
\eqref{eq:intro_active_stress} for the active strain and active stress
formulation respectively, are spatially resolved on the
LV free wall (LVFW), the RV free wall (RVFW) and the septum. With so
few control parameters, issues related to conflicting objectives, i.e
that is is not possible to minimize both the strain and volume, is
evident. However, with a low dimensional parameter space, questions
regarding identifiability and uniqueness of the estimated parameters
are easier to answer. In this study we also perform a validation of
the model, by comparing simulated and measured longitudinal strain
which is not used in the optimization.

@@

\subsection{Paper 4: Assesment region regional myocardial work using
  adjoint-based data assimilation}

@@

\subsection{Talks}
\begin{itemize}
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Hans
    Henrik Odland, Marie Rognes, and Samuel T. Wall. ``Patient
    Constrained Ventricular Stress Mapping'',
    Conference Presentation at MALT 2015,  Lugano, Switzerland (2015).
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Marie
    Rognes, and Samuel T. Wall. ``Personalization of a Cardiac
    Compuational Model using Clinical Measurements'', Conference
    Presentation at 28th Nordic Seminar on Computational
    Mechanics. Vol. 28. Tallin, Estonia, (2015).
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Marie
    Rognes, and Samuel T. Wall. ``Optimization of a Spatially Varying
    Cardiac Contraction parameter using the Adjoint Method'',
    Conference Presentation at FEniCS 16, Oslo, Norway,(2016).
  \item Finsberg, Henrik N., Gabriel Balaban, Joakim Sundnes, Hans
    Henrik Odland, Marie Rognes, and Samuel T. Wall. ``Personalized
    Cardiac Mechanical Model using a High Resolution Contraction Field
    '',  Conference Presentation at VPH16 Translating VPH to the
    Clinic,  Amsterdam, Netherlands (2016).
\end{itemize}

@@

\newpage
\section{Closing remarks and future directions}

Although we have shown in this thesis that
adjoint-based data assimilation is a powerful technique that opens new
possibilities in terms of patient specific mechanical simulations, many questions
still has to be answered before we can fully embrace the output of
such simulations.

@@

A rule of thump is that \emph{the spatial resolution of the parameters should be
  reflected in the spatial resolution of the observations}. This means
that if one is trying to fit data that are spatially resolved at some level,
then choosing parameters that are resolved at a finer level should be
done with caution. Regarding both paper 1, 2 and 4 we see that the
spatial resolution chosen was at a much finer level than the input
data. In this case, regularization techniques was used to restrict the
parameter space.

>>>>

In this paper we develop and test the pipeline for constructing a
patient specific mechanical simulation of a patient's heart, based on
clinical measurements and adjoint-based data assimilation
techniques.

@@

In this paper we extend the work in the two previous paper to
bi-ventricular geometries, and use it to estimate patient-specific
myofiber stress and indices of contractility. Furthermore, we investigate
the sensitivity of these computed features the helical fiber angle and
the choice of active modeling framework.

In this work we estimate the linear isotopic material parameter in
\eqref{eq:holzapel_trans}, spatially resolved on the LV and RV, by
minimizing the error in end-diastolic LV and RV cavity volumes. We also
apply a simplified algorithm for estimating in unloaded geometry, in
order to overcome issues related to estimating an unloaded
BiV-geometry, which is not well-posed. The amount of active
contraction is estimated during the active phase, by minimizing
simulated and measured volumes and circumferential strain. The active
control parameter, which are $\gamma$ in
\eqref{eq:intro_active_strain_Fa_gjerald}  and $T_a$ in
\eqref{eq:intro_active_stress} for the active strain and active stress
formulation respectively, are spatially resolved on the
LV free wall (LVFW), the RV free wall (RVFW) and the septum. With so
few control parameters, issues related to conflicting objectives, i.e
that is is not possible to minimize both the strain and volume, is
evident. However, with a low dimensional parameter space, questions
regarding identifiability and uniqueness of the estimated parameters
are easier to answer. In this study we also perform a validation of
the model, by comparing simulated and measured longitudinal strain
which is not used in the optimization.

@@

\subsection{Paper 4: Assessment of regional myocardial work from a
  patient-specific cardiac mechanics model}

\begin{figure}[htbp]
  \centering
    \includegraphics[width=0.7\textwidth]{chapters/introduction/figures/paper4}
\caption{Summary figure for paper 4}
\label{fig:paper3}
\end{figure}

In this paper we investigate if it is possible to assess regional
myocardial work using data assimilation and finite element modeling.
The amount of work performed by the left ventricle equals to the
amount to work needed to eject blood against the intraventricular
pressure. Effectively, a global measure of work is therefore the same
as the area of the pressure volume (PV) loop which is most commonly refered
to as stroke work.  However, stroke work is a global measure which
does not reflect whether some regions in the ventricle work harder than
others to compensate for regional dysfunction, in order to maintain the
stroke work needed to meet the body's demand. Therefore a regional
measure of work would be helpful in assessment of patients with
regional dysfunction, and preferable we would have a regional meausure
that summed up to the actual stroke work. The area of the
pressure-strain loops, where strain here refers to as regional
longitudinal strain, has been proposed as a meausure regional
myocardial work, and indices based on these work calculations have
shown to be a better predictor response to cardiac resynchronization
therapy, a treatment given to patients with regional dysfunction.

To check the validity of these work calculations we estimate work
using well known laws from continuum mechanics in a personalized
cardiac mechanics model, similar to the models in the previous papers.
These work calculations are compared to estimates using the
pressure-strain loops approach, and to the ``gold-standard'' PV area.
Our estimates of global mechanical work are in good agreements with
estimated PV area, while the pressure-strain loops quantitatively
underestimates the total mechanical work. However, relative measures
of regional efficiency seems to be independent of how work is
computed, and yields the same results in terms of assesment of regional
dysfunction. Our finding therefore suggest that regional myocardial
work can be assessed using patient-specific finite element modeling,
while regional dysfunction in terms of indices of efficiency are
independent of how work is computed.

In this study we also perform a convergence analysis of the algorithm for
determining the unloaded, zero-pressure geometry as well as the
algorithm for joint estimation of material parameters and unloaded
geometry.

@@

\subsection{Talks}
\begin{itemize}
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Hans
    Henrik Odland, Marie Rognes, and Samuel T. Wall. ``Patient
    Constrained Ventricular Stress Mapping'',
    Conference Presentation at MALT 2015,  Lugano, Switzerland (2015).
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Marie
    Rognes, and Samuel T. Wall. ``Personalization of a Cardiac
    Computational Model using Clinical Measurements'', Conference
    Presentation at 28th Nordic Seminar on Computational
    Mechanics. Vol. 28. Tallin, Estonia, (2015).
  \item Henrik Finsberg, Gabriel Balaban, Joakim Sundnes, Marie
    Rognes, and Samuel T. Wall. ``Optimization of a Spatially Varying
    Cardiac Contraction parameter using the Adjoint Method'',
    Conference Presentation at FEniCS 16, Oslo, Norway,(2016).
  \item Finsberg, Henrik N., Gabriel Balaban, Joakim Sundnes, Hans
    Henrik Odland, Marie Rognes, and Samuel T. Wall. ``Personalized
    Cardiac Mechanical Model using a High Resolution Contraction Field
    '',  Conference Presentation at VPH16 Translating VPH to the
    Clinic,  Amsterdam, Netherlands (2016).
\end{itemize}

@@

\newpage

% \section{Aim of thesis}
% Utilizing the power of computational modeling and advanced numerical
% algorithms to target the problem of personalized cardiac mechanics
% modeling has been the primary motivation for the work conducted in 
% this thesis. In particular, we wanted to explore the use of 
% variational data-assimilation techniques to estimate model parameters
% based on minimization of mismatch between observed, clinical data and
% simulated data from a cardiac mechanical model. 
% The developed framework in this thesis could be applied directly to estimate high
% dimensional model parameters without any manual tuning of
% parameters. The aim has been to develop a framework for automatic
% generation of patient-specific simulations of the heart, by fitting model-parameters to
% potentially ``any'' type of clinical data. In particular,
% incorporating regional motion data to estimate spatially resolved
% parameters has been an essential contribution in this thesis.
% This type of automated simulation pipeline could potentially be integrated as a part of a
% clinical diagnostic toolbox and be used to guide clinicians in the
% decision-making and treatment planning. In particular, such models can
% be used to assess information that is difficult or even impossible to
% measure in the human heart, such as indices of contractility,
% myocardial work and myofiber stress. While the first part of the
% thesis (Paper 1) is concerned with developing a framework for
% constraining a mechanics model to clinical data, the second part
% (Paper 2, 3 and 4) evaluates potentially use biomarkers extracted from
% the personalized mechanics model. Future work should be geared towards
% validation of such models, in order to unleash its potential in the
% clinic.

\section{Closing remarks and future directions}

Although we have shown in this thesis that the techniques developed in
this thesis are powerful, and opens up new possibilities in terms of
patient specific mechanical simulations, many questions still has to
be answered before we can fully embrace the output of such simulations.

@@

A rule of thump is that \emph{the spatial resolution of the parameters should be
  reflected in the spatial resolution of the observations}. This means
that if one is trying to fit data that are spatially resolved at some level,
then choosing parameters that are resolved at a finer level should be
done with caution. Regarding both paper 1, 2 and 4 we see that the
spatial resolution chosen was at a much finer level than the input
data. In this case, regularization techniques was used to restrict the
parameter space, by assuming that the solution you are seeking has
some particular properties, e.g smoothness.

### 2017-09-26 chapters/paper1/article.tex
Add latest
Write summary. Only missing to write about fenics and summary of  the unpublished manuscripts.
Added paper 3 and update introduction

\chapter{Paper 1: \\High Resolution Data Assimilation of Cardiac
  Mechanics}

% \pagebreak
\clearpage

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

% \begin{center}
G. Balaban\footnotemark,
H. Finsberg\footnotemark[\value{footnote}], 
H. Odland, M. E. Rognes, \\
S. Ross, J. Sundnes, and
S. Wall
% \end{center}

@@

\clearpage
\section{Introduction}
Computational models of cardiac mechanics, personalized to the level of the individual 
through the use of clinical imaging, have potential to be a powerful aid in the diagnosis and 
treatment of cardiac disease.  By relating image-based data to fundamental physical 
processes, models can give additional insight into the function or dysfunction of the 
individual's heart, beyond what can be directly measured or observed in the images. 
This is becoming more important as the resolution and accuracy of clinical imaging continues to 
improve.  This increasingly detailed data combined with biophysical models has promise 
in analysis of regionally and temporally resolved differences in the mechanics of the heart, 
important in diseases such as heart failure and the application of cardiac resynchronization therapy.

@@

\begin{table}
\caption{Sensitivity of optimal $a$ value to choice of spring constant $k$.}
\begin{tabular}{|l|llllllllllll|}
\hline
      $k$ & $10^{-8}$ & $10^{-7}$& $10^{-6}$&  $10^{-5}$  & $10^{-4}$ & $10^{-3}$ & 0.01 & 0.1 & 1 & 10 & 100 & $\infty$ \\
\hline
      $a$ & 0.875   & 0.875  & 0.875  & 0.875 & 0.873& 0.873 & 0.849 & 0.684 & 0.435 & 0.375& 0.366 & 0.365 \\
\hline
\end{tabular}
\label{tab:elast_sense}
\end{table}

@@

% \FloatBarrier
\pagebreak

\bibliographystyle{plain}
\bibliography{chapters/paper1/bibliography}

>>>>

\chapter{Paper 1}
{\Huge \textbf{High Resolution Data Assimilation of Cardiac
  Mechanics}}

% \pagebreak
% \clearpage
\newpage
\renewcommand{\thefootnote}{\fnsymbol{footnote}}

@@

% \begin{center}
Gabriel Balaban$^{1,2,5}$\footnotemark[\value{footnote}],
Henrik Finsberg$^{1,2,5}$\footnotemark,
Hans Henrik Odland$^{4,7}$,
M. E. Rognes$^{1,3}$,
Stian Ross$^{4,4}$,
Joakim Sundnes$^{1,2,5}$, and
Samuel Wall$^{1,5,6}$

\footnotesize
\begin{enumerate}[itemsep=-2mm]
\item{Simula Research Laboratory, Lysaker, Norway}
\item{Department of Informatics, University of Oslo, Oslo, Norway}
\item{Department of Mathematics, University of Oslo, Oslo, Norway}
\item{Faculty of Medicine, University of Oslo, Oslo, Norway}
\item{Center for Cardiological Innovation, Oslo, Norway}
\item{Department of Mathematical Science and Technology, NMBU, \r{A}s,
    Norway}
\item{Department of Pediatrics, Oslo University Hospital, Oslo, Norway}
\end{enumerate}
\normalsize

@@

% \clearpage
\newpage
\section{Introduction}
Computational models of cardiac mechanics, personalized to the level of the individual 
through the use of clinical imaging, have potential to be a powerful aid in the diagnosis and 
treatment of cardiac disease.  By relating image-based data to fundamental physical 
processes, models can give additional insight into the function or dysfunction of the 
individual's heart, beyond what can be directly measured or observed in the images. 
This is becoming more important as the resolution and accuracy of clinical imaging continues to 
improve.  This increasingly detailed data combined with biophysical models has promise 
in analysis of regionally and temporally resolved differences in the mechanics of the heart, 
important in diseases such as heart failure and the application of cardiac resynchronization therapy.

@@

\begin{table}
\caption{Sensitivity of optimal $a$ value to choice of spring constant
  $k$.}
\scalebox{0.8}{
\begin{tabular}{|l|llllllllllll|}
\hline
      $k$ & $10^{-8}$ & $10^{-7}$& $10^{-6}$&  $10^{-5}$  & $10^{-4}$ & $10^{-3}$ & 0.01 & 0.1 & 1 & 10 & 100 & $\infty$ \\
\hline
      $a$ & 0.875   & 0.875  & 0.875  & 0.875 & 0.873& 0.873 & 0.849 & 0.684 & 0.435 & 0.375& 0.366 & 0.365 \\
\hline
\end{tabular}}
\label{tab:elast_sense}
\end{table}

@@

% \FloatBarrier
% \pagebreak
\newpage
\bibliographystyle{plain}
\bibliography{chapters/paper1/bibliography}

### 2017-09-26 chapters/paper2/main.tex
Add latest
Added paper 3 and update introduction

\chapter{Paper 2: \\Estimating cardiac contraction through high resolution data
  assimilation of a personalized mechanical model}

\clearpage

\section*{Estimating cardiac contraction through high resolution data
  assimilation of a personalized mechanical model}

\begin{center}
  Henrik Finsberg
  Gabriel Balaban,
  Stian Ross,
  Trine H\r{a}land,
  Hans Henrik Odland, 
  Joakim Sundnes, and
  Samuel Wall
\end{center}

@@

\clearpage
%% main text
\section{Introduction}
\label{sec:introduction}

@@

\pagebreak

\bibliographystyle{plain}
\bibliography{chapters/paper2/bibliography}

>>>>

\chapter{Paper 2}
{\Huge \textbf{Estimating cardiac contraction through high resolution data
  assimilation of a personalized mechanical model}}

% \clearpage
\newpage
\section*{Estimating cardiac contraction through high resolution data
  assimilation of a personalized mechanical model}

% \begin{center}
  Henrik Finsberg$^{1,2,4}$,
  Gabriel Balaban$^6$,
  Stian Ross$^{3,4}$,
  Trine H\r{a}land$^{3,4}$,
  Hans Henrik Odland$^{3,4}$,
  Joakim Sundnes$^{1,2,4}$, and
  Samuel Wall$^{1,3,5}$
% \end{center}

\footnotesize
\begin{enumerate}[itemsep=-2mm]
\item{Simula Research Laboratory, Lysaker, Norway}
\item{Department of Informatics, University of Oslo, Oslo, Norway}

\item{Faculty of Medicine, University of Oslo, Oslo, Norway}
\item{Center for Cardiological Innovation, Oslo, Norway}
\item{Department of Mathematical Science and Technology, NMBU, \r{A}s,
    Norway}
\item{Department of Imaging Sciences and Bioengineering,
    KCL, London, England}
\end{enumerate}
\normalsize

@@

% \clearpage
\newpage
%% main text
\section{Introduction}
\label{sec:introduction}

@@

% \pagebreak
\newpage
\bibliographystyle{plain}
\bibliography{chapters/paper2/bibliography}
